\chapter{Build and Command Line}

\section{Build Requirements}

The project uses CMake and C++20. Optional runtime functionality is enabled when dependencies are available:

\begin{itemize}
  \item SDL3 enables the runtime window and renderer;
  \item Lua 5.4 enables game scripts;
  \item libmpg123 enables MP3 audio playback.
\end{itemize}

Missing optional dependencies do not prevent the project from compiling. Instead, the corresponding runtime feature is compiled out or skipped.

\section{Standard Commands}

The expected development commands are:

\begin{lstlisting}[language=bash]
cmake --preset linux-debug
cmake --build --preset linux-debug
ctest --preset linux-debug
./build/linux-debug/demi --help
./build/linux-debug/demi validate examples/minimal_2d/demi.project.json
\end{lstlisting}

The runtime can be launched with:

\begin{lstlisting}[language=bash]
./build/linux-debug/demi run --project examples/minimal_2d/demi.project.json
\end{lstlisting}

For headless smoke testing, SDL can use its dummy video driver:

\begin{lstlisting}[language=bash]
SDL_VIDEODRIVER=dummy ./build/linux-debug/demi run --project examples/minimal_2d/demi.project.json --frames 5
\end{lstlisting}

\section{Build Targets}

The current CMake configuration defines these major targets:

\begin{longtable}{>{\ttfamily}p{0.28\textwidth}p{0.62\textwidth}}
demi-core & Core validation, filesystem, diagnostics, and asset registry code. \\
demi-runtime-lib & Runtime systems including scene loading, rendering, physics, audio, and Lua hosting. \\
demi & Main command line executable. \\
demi-runtime & Runtime-specific executable entry point. \\
demi-editor & Editor scaffold. \\
demi-smoke-tests & Core smoke tests. \\
demi-physics2d-tests & Physics regression tests. \\
\end{longtable}
