\chapter{Minimal 2D Example}

The \filepath{examples/minimal\_2d} project is the main gameplay example. It demonstrates project loading, scene data, generated gameplay objects, Lua scripts, physics, audio, saves, and the HUD menu system.

\section{Project Contents}

Important files include:

\begin{longtable}{>{\ttfamily}p{0.42\textwidth}p{0.46\textwidth}}
demi.project.json & Project entry point. \\
scenes/main.scene.json & Level 1 scene and entity definitions. \\
scenes/spiral.scene.json & Level 2 scene: a procedurally generated upward spiral climb. \\
scenes/main.hud.json & Declarative main menu, level select, options UI, and points HUD. \\
scripts/game.lua & Shared game controller: points, coins, platform tracking, and level dispatch. \\
scripts/level\_main.lua & Level 1 generation and camera (horizontal platforms). \\
scripts/level\_spiral.lua & Level 2 generation and camera (upward spiral climb). \\
scripts/game\_state.lua & Shared state across scene switches (level, auto\_start, fall floor). \\
scripts/player.lua & Player update entry point. \\
scripts/player\_platformer.lua & Platformer movement and respawn helpers. \\
scripts/player\_slingshot.lua & Slingshot aim, recharge, launch, and trajectory preview. \\
scripts/main\_menu.lua & Menu state controller, level select, and settings persistence. \\
scripts/menu\_button.lua & HUD button event dispatcher. \\
saves/settings.save.json & Generated or updated settings save file. \\
\end{longtable}

\section{Levels and Scene Switching}

The example ships two levels. The main menu has a Level Select screen showing one card per level (number, name, description, a mini preview, and a PLAY button). Selecting a level calls \code{Scene.load} with the target scene id; the runtime rebuilds the world and scripts between frames (see the Scene Switching section in the Lua API chapter). A shared state module carries an \code{auto\_start} flag from the menu button into the new scene's \code{on\_start} so the selected level begins immediately. On the Level Select screen, keyboard \code{1}/\code{2} pick a level and \code{Escape} returns to the main menu.

\begin{itemize}
  \item \textbf{Level 1 -- Orbit Run} (\filepath{scenes/main.scene.json}): the original deterministic horizontal slingshot platformer. Platforms generate ahead of the camera along the x axis.
  \item \textbf{Level 2 -- Spiral Climb} (\filepath{scenes/spiral.scene.json}): a procedurally generated upward spiral. Platforms wind around a center point at a fixed angle step with a slowly growing radius, ascending by a fixed height per step. The camera follows the player on both axes. The fall floor is the same fixed value as Level 1, so a missed jump respawns at the start just like the main level.
\end{itemize}

\section{Gameplay}

The example is a 2D slingshot platformer. The player can move, jump, bounce on platforms, collect coins, and use a slingshot mechanic. Platform generation is deterministic and happens ahead of the camera.

Recent gameplay behavior includes:

\begin{itemize}
  \item slingshot can recharge after touching or bouncing off platforms;
  \item mid-air slingshot preserves bounce velocity;
  \item trajectory preview uses snapshotted velocity for a stable prediction line;
  \item platform edge collision avoids teleporting across the platform.
\end{itemize}

\section{Menu}

The main menu is authored in \filepath{main.hud.json}. Lua controls group visibility and dynamic settings. The menu includes:

\begin{itemize}
  \item cosmic themed start screen;
  \item Start Game, Options, and Quit buttons;
  \item Sound and Video tabs;
  \item master volume controls;
  \item window mode dropdown for windowed, borderless, and fullscreen;
  \item persistent settings through the \code{Save} service.
\end{itemize}
